---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# labNorm

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/labNorm)](https://CRAN.R-project.org/package=labNorm)
<!-- badges: end -->

`labNorm` provides functions for normalizing standard laboratory measurements (e.g. hemoglobin, cholesterol levels) according to age and gender, based on the algorithms described in ["Personalized lab test models to quantify disease potentials in healthy individuals"](https://doi.org/10.1038/s41591-021-01468-6). 

Allows users to easily obtain normalized values for standard lab results, and to visualize their distributions. For more information go to: https://tanaylab.weizmann.ac.il/labs/

## Installation

You can install the development version of labNorm like so:

``` r
retmotes::install_github("tanaylab/labNorm")
```

## Example

Normalize Hemoglobin values to age and sex: 

```{r}
library(labNorm)
hemoglobin_data$quantile <- ln_normalize(
    hemoglobin_data$value,
    hemoglobin_data$age,
    hemoglobin_data$sex,
    "Hemoglobin"
)
head(hemoglobin_data)
```

Plot the quantiles vs values for age 50-60: 

```{r}
library(ggplot2)
library(dplyr)
hemoglobin_data %>%
    filter(age >= 50 & age <= 60) %>%
    ggplot(aes(x = value, y = quantile, color = sex)) +
    geom_point() +
    theme_classic()
```
